---
title: "Progress Report 2"
Date: "12/June/2019"
GitHub: "https://github.com/klbrundage/LLO8200-Assignments"
output: html_document
fontsize: 12pt
geometry: margin=.5
autosize: yes
Author: "Kelley Brundage"
---

```{r setup, include=FALSE}
##This code allows the Knit function to still work even with errors 
knitr::opts_chunk$set(echo=TRUE,error=TRUE)
```

```{r global_options, include = FALSE}
##This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.
knitr::opts_chunk$set(message=FALSE, 
tidy.opts=list(width.cutoff=60)) 

my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}
```

```{r eval=F, tidy=T}
##The `eval=FALSE` option just displays the R code (and does not run it), `tidy=TRUE` wraps long code so it does not run off the page.

##Below are the standard set of setup commands and common libraries that may be needed when running data analysis.
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(dplyr))
suppressMessages(library(evaluate))
suppressMessages(library(forcats))
suppressMessages(library(formatR))
suppressMessages(library(ggplot2))
suppressMessages(library(haven))
suppressMessages(library(knitr))
suppressMessages(library(ModelMetrics))
suppressMessages(library(modelr))
suppressMessages(library(readxl))
suppressMessages(library(RColorBrewer))
suppressMessages(library(tibble))
suppressMessages(library(tidyverse))
```

```{r Import World Ranking Datasets}
##The code below will upload the World University Ranking Datasets obtained from Kaggle.com

##The next set of files are the three World Ranking system files
library(readxl)
cwurData <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/cwurData.xlsx")
View(cwurData)

library(readxl)
shanghaiData <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/shanghaiData.xlsx")
View(shanghaiData)

library(readxl)
timesData <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/timesData.xlsx")
View(timesData)

##The three datasets below are the additional datasets provided that connect to the World University Ranking datasets
library(readxl)
education_expenditure_supplementary_data <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/education_expenditure_supplementary_data.xlsx")
View(education_expenditure_supplementary_data)

library(readxl)
educational_attainment_supplementary_data <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/educational_attainment_supplementary_data.xlsx")
View(educational_attainment_supplementary_data)

##The file below lists the name of the institution and the Country the institution is located in
library(readxl)
school_and_country_table <- read_excel("~/School/EdD program - Vanderbilt/01 - Classes/LLO 8200 - Intro to Data Science/Final Project/datafiles/school_and_country_table.xlsx")
View(school_and_country_table)
```

# Progress Report 2:  World University Rankings

##*Synopsis of Problem & Approach*
Compare the three global ranking systems to the amount of faculty, publications/research at each institution per ranking system by approaching each dataset with an analytical and statistical viewpoint. 

Analyze the dataset specific to the area of research/academic and if common challenges that exist with all ranking systems exist.  Problems such as making the correction for institutional size, differences between average and extreme, defining the institutions, measurement of time frame, credit allocation, excellency factors as well as adjustment for scientific fields or types of research.

##*Columns in each dataset that will be used to compare and analysze the ranking system*

CWUR Dataset Fields:
world_rank; institution; country; national_rank; publications 
#Quality of Faculty, measured by the number of academics who have won major international awards, prizes, and medals (15%)

#Quality Publications, measured by the number of research papers appearing in top-tier journals (15%) 

Shanghai Dataset Fields:
word_rank; university_name; national_rank; pub

Times Dataset Fields:
world_rank; university; country; research
#Research (volume, income, and reputation)

#Citations (research influence)

# **Extensive Investigation of Dataset**

Investigate data: distribution of data, correlations, associations, and predictive potential to solve your proposed problem

```{r Regression for CWUR Data - Publications}
cwurData%>%summarize(mean(publications,na.rm=TRUE))
  #na.rm=T specifies what to do with any missing data in the dataset

cwurData<-cwurData%>%mutate(publications_p=percent_rank(publications)*100)
  #default would give 0-1 so *by 100 and it will make it easier to understand the publication number in the CWUR dataset

cwurData%>%summarize(mean(publications_p,na.rm=TRUE)) 
  #gives the mean of the publications in the CWUR dataset
```

```{r Regression for Shanghai Data - Publications}
shanghaiData%>%summarize(mean(pub,na.rm=TRUE))
  #na.rm=T specifies what to do with any missing data in the dataset

shanghaiData<-shanghaiData%>%mutate(pub_p=percent_rank(pub)*100)
  #default would give 0-1 so *by 100 and it will make it easier to understand the publication number in the Shanghai Dataset

shanghaiData%>%summarize(mean(pub_p,na.rm=TRUE)) 
  #gives the mean of the publications in the Shanghai dataset
```


```{r Regression for Times Data - Publications}
timesData%>%summarize(mean(citations,na.rm=TRUE))
  #na.rm=T specifies what to do with any missing data in the dataset

timesData<-timesData%>%mutate(citations_p=percent_rank(citations)*100)
  #default would give 0-1 so *by 100 and it will make it easier to understand the publication number in the Times Dataset

timesData%>%summarize(mean(citations_p,na.rm=TRUE)) 
  #gives the mean of the publications in the Times dataset
```

```{r Histogram of Publications Regression for CWUR}
gc1 <- ggplot(cwurData,aes(x=publications_p))
gc1 <- gc1+geom_density(binwidth = 1,fill="darkgreen")
gc1
```

```{r Histogram of Publications Regression for Shanghai}
gs1 <- ggplot(shanghaiData,aes(x=pub_p))
gs1 <- gs1+geom_density(binwidth = 1,fill="purple")
gs1
```

```{r Histogram of Publications Regression for Times}
gt1 <- ggplot(timesData,aes(x=citations_p))
gt1 <- gt1+geom_density(binwidth = 1,fill="darkblue")
gt1
```

##*World Rankings Dataset Predictions*

##CWUR Dataset Prediction
```{r CWUR Publications Prediction}
cpub1<-lm(publications~world_rank,data=cwurData) 
  #publications ~(as a function of) world_rank, data=dataset(cwur)

#outcome (publications) on left, predictor (world rank) on right 
summary(cpub1)#shows the results of the regression
```

RESULTS:
if world rank is 0 then intercept(publications) are predicted to be 36 (Est. Std)
as world rank increases (every 1 unit change) publicatinos are predicted to increase by .9 points (Est. Std)

Reject the Null Hypothesis that the coefficient is zero

Residual Standard Error (RMSE): 116.9 on df(2198)

```{r CWUR Coefficient Data}
confint(cpub1)
#This code only shows the coefficient Data
```

```{r Point Plot of CWUR Publications by World Rank}
gp2 <- ggplot(cwurData,aes(x=world_rank,y=publications))+ 
      #x is the world rank and y is the publicatons
  geom_point(shape=1)+#specifies the points
  geom_smooth(method = lm)
gp2 <- gp2+ggtitle("Publications by World Rank in CWUR Dataset")
gp2
```


```{r RMSE for CWUR publications}
cwurData<- cwurData%>%add_predictions(cpub1)%>%rename(predc1=pred)
  #predict using data in memory
  
rmse_cpub1<-modelr::rmse(cpub1,cwurData);rmse_cpub1
  #on average we are off by 117 points
```

##Shanghai Dataset Prediction
```{r Shanghai Publications Prediction}
spub1<-lm(pub~world_rank,data=shanghaiData) 
  #publications ~(as a function of) world_rank, data=dataset(shanghai)

#outcome (publications) on left, predictor (world rank) on right 
summary(spub1)#shows the results of the regression
```

```{r Shanghai Coefficient Data}
confint(spub1)
#This code only shows the coefficient Data
```

```{r Point Plot of Shanghai Publications by World Rank}
gs2 <- ggplot(shanghaiData,aes(x=world_rank,y=pub))+ 
      #x is the world rank and y is the publicatons
  geom_point(shape=1)+#specifies the points
  geom_smooth(method = lm)
gs2 <- gs2+ggtitle("Publications by World Rank in Shanghai Dataset")
gs2
```

```{r RMSE for Shanghai publications}
shanghaiData<- shanghaiData%>%add_predictions(spub1)%>%rename(preds1=pred)
  #predict using data in memory
  
rmse_spub1<-modelr::rmse(spub1,shanghaiData);rmse_spub1
  #on average we are off by 7 points
```

##Times Dataset Prediction
```{r Times Publications Prediction}
tpub1<-lm(citations~world_rank,data=timesData) 
  #publications ~(as a function of) world_rank, data=dataset(cwur)

#outcome (publications) on left, predictor (world rank) on right 
summary(tpub1)#shows the results of the regression
```

```{r Times Coefficient Data}
confint(tpub1)
#This code only shows the times coefficient Data
```

```{r Point Plot of Times Publications by World Rank}
gt2 <- ggplot(timesData,aes(x=world_rank,y=citations))+ 
      #x is the world rank and y is the publicatons
  geom_point(shape=1)+#specifies the points
  geom_smooth(method = lm)
gt2 <- gt2+ggtitle("Citations by World Rank in Times Dataset")
gt2
```

```{r RMSE for Times publications}
timesData<- timesData%>%add_predictions(tpub1)%>%rename(predt1=pred)
  #predict using data in memory
  
rmse_tpub1<-modelr::rmse(tpub1,timesData);rmse_tpub1
  #on average we are off by 12 points
```


```{r Prediction of the Publications under the 3 World Ranking Datasets}
#Prediction from the Publications data under the 3 world ranking datasets
rmse_cpub1
rmse_spub1
rmse_tpub1
```

